---
title: "experiment-1a_dutch"
output: html_document
date: "2023-01-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up

```{r setup}

###### folders ###########

#current folder (first go to session -> set working directory -> to source file location)
parentfolder <- dirname(getwd())
print(parentfolder)
#list the folders
datasets    <- paste0(parentfolder, '/Data/')
folderdata  <- paste0(datasets, '/raw_data/')

print(folderdata)

##### packages ######
library(readr)
library(R.utils)
library(dplyr)
library(tidyr)
library(readxl)

```

How many participants finished until the end?
```{r participants}

# info from Pavlovia
## 237 completed
## 109 aborted

list_data <- list.files(paste0(folderdata), pattern = ".csv")
length(list_data) # 264 csv files in total, the empty ones or aborted early before trial round deleted manually


```

One df
```{r making one df}

## Read the first csv file as a start for combining all individual dataframes
data = read.csv(paste0(folderdata, "/001324_flesh_expressibility_dutch.csv"))

## Write another column 'file' with a filename
data$file = '001324_flesh_expressibility_dutch.csv'

dirs = list.files(paste0(folderdata), pattern = ".csv")

##### Combine dataframes #####

for (ifile in dirs){
  ## Set the path according to the mother directory
  newdata = read.csv(paste0(folderdata, "/", ifile))
  newdata$file = ifile
  
  data=plyr::rbind.fill(data,newdata)
    
  }
  

### remove duplicate rows
data <- unique(data)

```


Filtering data frame
```{r filtering data frame}

# Deselect unneeded variables:

# variables that has only NAs
data <- select(data, -c(natlang4, natlang5, natlang6))
data <- select(data, -c(lang6, lang7, lang8, proficiency6, proficiency7, proficiency8))

# Psychopy metadata
data <- select(data, -c(psychopyVersion, OS, frameRate, expName, session, presentation.limit, date))
data <- select(data, -c(xRes, yRes))
# unnecesary timings etc
data <- select(data, -c(introduction_weiter.numClicks, introduction_weiter.timesOn, introduction_weiter.timesOff, mouse.x, mouse.y, mouse.leftButton, mouse.midButton,  mouse.rightButton, mouse.time, mouse.clicked_name, button_confirm.numClicks, button_confirm.timesOn, button_confirm.timesOff, button_exit.numClicks, button_exit.timesOn, button_exit.timesOff, practice_weiter.numClicks, practice_weiter.timesOn, practice_weiter.timesOff, pr_message_weiter.numClicks, pr_message_weiter.timesOn, pr_message_weiter.timesOff, pr_remember_weiter.numClicks, pr_remember_weiter.timesOn, pr_remember_weiter.timesOff,  button_weiter.numClicks, button_weiter.timesOn, button_weiter.timesOff, pr_image, start_weiter.numClicks, trials.order))

# unnecesary timings vol 2
data <- select(data, -c(welcome_weiter.numClicks, welcome_weiter.timesOn, welcome_weiter.timesOff, radio.for.required.checkboxes, sensorypref.., practice_trials.thisRepN, practice_trials.thisTrialN, practice_trials.thisN, practice_trials.thisIndex, practice_trials.ran, start_weiter.timesOff, message_weiter.numClicks, message_weiter.timesOn, message_weiter.timesOff, trial_remember_weiter.numClicks, trial_remember_weiter.timesOn, trial_remember_weiter.timesOff, Count, Shelf.time, trial_button_weiter.numClicks, trial_button_weiter.timesOn, trial_button_weiter.timesOff, trials.thisRepN, trials.thisTrialN, trials.thisN, trials.thisIndex, trials.ran, fordict, trial_img, end_experiment.numClicks, end_experiment.timesOff))

## delete practice round
data <- select(data, -c(pr_expressibility.response, pr_expressibility.rt, pr_word, pr_question))

## some more
data <- select(data, -c(blocks.thisRepN, blocks.thisTrialN, blocks.thisN, blocks.thisIndex, blocks.ran, blocks.order, trial_list, task, mod_image, remember))

### remove duplicate rows
data <- unique(data)


## change name of columns
colnames(data)[26]  <- "trial_start_time" 
colnames(data)[27]  <- "expressibility" 
colnames(data)[28]  <- "response_time"
colnames(data)[30]  <- "modality"
colnames(data)[34]  <- "end_time"


```

Detection of problems
```{r any problems}

# signs of test
unique(data$natlang1)
unique(data$lang1) 
unique(data$problem_text)
unique(data$comment_text)
unique(data$exp_ID)

# 
# ## find the corresponding participants
row_index <- which(data$comment_text == "enkele keer verschoof het balletje op de balk nadat ik op volgende had gedrukt. ")
which <- data[row_index, "exp_ID"]
unique(which)

# delete test trials
data <- data %>% filter(exp_ID != 22894 & exp_ID != 221539 & exp_ID != 410740 & exp_ID != 590154 & exp_ID != 682029 & exp_ID != 70523 & exp_ID != 'NA')

# delete pilot 
data <- data %>% filter(exp_ID != 232400 & exp_ID != 254289 & exp_ID != 682241 & exp_ID != 148688 & exp_ID != 582962 & exp_ID != 'wimpouw' & exp_ID != '%SURVEY_CODE%')

# delete ppcns using phone/ipad
data <- data %>% filter(exp_ID != 789985 & exp_ID != 968829)

# problems
## 141475
# er stond telkens bij de herinnering niet hetzelfde als boven de vraag, want daar stond zonder taal
## 417863
# Door de tijd die het koste om de woorden te laden, vergat ik een beetje wat ik wel en niet mocht doen.
## 683060
# Het duurde erg lang tot de volgende vraag kwam, heb af en toe echt 30 seconden moeten wachten!
## 723521
# bij sommige vragen verschoof de stip mee met de muis voordat ik op verder drukte
## 871335
# , er stond geen duidelijke minimum leeftijd aangegeven maar de leeftijden startte vanaf 18 jaar 


# comments
## 871335
# Ik heb het feit dat NGT bestaat even achterwege gelaten
## 832980
# enkele keer verschoof het balletje op de balk nadat ik op volgende had gedrukt.


# exp_ID to delete, bugs

data <- data %>% 
  filter(exp_ID != 738964 & exp_ID != 439659 & exp_ID != 126241 & exp_ID != 490543 & exp_ID != 914515 & exp_ID != 510856 & exp_ID != 895055 & exp_ID != 372271 & exp_ID != 426597 & exp_ID != 126241 & exp_ID != 639780)


## we can delete problem columns now
data <- select(data, -c(problem, problem_text))

length(unique(data$exp_ID)) ## 228 unique participants (not all completed)

```

We need to convert time info to numeric and distribute all over the df
```{r update some df things}

# times as numeric

data <- data %>% mutate(trial_start_time = parse_number(data$trial_start_time))
data <- data %>% mutate(end_time = parse_number(data$end_time))

# times everywhere in the column 

data <- data %>% group_by(exp_ID) %>% 
  fill(trial_start_time) %>% 
  fill(trial_start_time, .direction = "up") %>% 
  ungroup()

data <- data %>% group_by(exp_ID) %>% 
  fill(end_time) %>% 
  fill(end_time, .direction = "up") %>% 
  ungroup()

## delete rows where expressibility is NA

data <- data %>% filter(!is.na(expressibility))

## add duration column

data$duration <- NA 
data$duration <- data$end_time - data$trial_start_time


```

Renaming columns and correcting language information
```{r naming corrections}

## rename modality column

data$modality <- ifelse(data$modality == "question_vocal.png", "vocal", 
                      ifelse(data$modality == "question_multimodal.png", "multimodal",
                             ifelse(data$modality == "question_gesture.png", "gesture", NA)))


# natlang1

# ## find the corresponding participants
row_index <- which(data$lang1 == "engels, Duits en een beetje Frans")
which <- data[row_index, "exp_ID"]
unique(which)


data$natlang1 <- "nederlands"

# correct pcn 728515
data$natlang1[data$exp_ID == 728515] <- "nederlands"
data$natlang2[data$exp_ID == 728515] <- "chinees"

# natlang2
unique(data$natlang2)

# people mentioning dialect
## 257220 Limburgs
## 567345 Dialect
data$dialect <- NA
data$natlang2[data$exp_ID == 257220] <- NA
data$dialect[data$exp_ID == 257220] <- "limburgs"
data$natlang2[data$exp_ID == 567345] <- NA
data$dialect[data$exp_ID == 567345] <- "dialect"

# uniforming
data$natlang2 <- gsub("Farsi ", 'farsi', data$natlang2)
data$natlang2 <- gsub("-", NA, data$natlang2)
data$natlang2 <- tolower(data$natlang2)

#natlang3
unique(data$natlang3)
data$natlang3 <- tolower(data$natlang3)

## Note, pcn 680990 have ovelap between natlangs and langs

unique(data$lang1) 
## Note, pcn 901632 overlap between natlang and lang

## correct participants
###398421
data$lang1[data$exp_ID == 398421] <- 'duits'
data$lang2[data$exp_ID == 398421] <- 'engels'
data$lang3[data$exp_ID == 398421] <- 'frans'
data$lang4[data$exp_ID == 257220] <- 'zweeds'

## 784091
data$lang1[data$exp_ID == 784091] <- 'engels'
data$lang2[data$exp_ID == 784091] <- 'duits'

##18102	
data$lang1[data$exp_ID == 18102] <- 'engels'
data$lang2[data$exp_ID == 18102] <- 'duits'

##115296
data$lang1[data$exp_ID == 115296] <- 'engels'
data$lang2[data$exp_ID == 115296] <- 'duits'
data$proficiency2[data$exp_ID == 115296] <- 2

##319675		
data$lang1[data$exp_ID == 319675] <- 'engels'
data$lang2[data$exp_ID == 319675] <- 'duits'
data$proficiency2[data$exp_ID == 319675] <- 2

##512122	
data$lang1[data$exp_ID == 512122] <- 'engels'
data$lang2[data$exp_ID == 512122] <- 'duits'

##552962
data$lang1[data$exp_ID == 552962] <- 'engels'
data$lang2[data$exp_ID == 552962] <- 'duits'

##882595
data$lang1[data$exp_ID == 882595] <- 'engels'
data$lang2[data$exp_ID == 882595] <- 'duits'
data$proficiency2[data$exp_ID == 882595] <- 2
data$lang3[data$exp_ID == 882595] <- 'frans'
data$proficiency2[data$exp_ID == 882595] <- 1 # Note, in answer 'een beetje'

##208529
data$lang1[data$exp_ID == 208529] <- 'engels'
data$lang3[data$exp_ID == 208529] <- 'duits'
data$proficiency3[data$exp_ID == 208529] <- 2

##625900
data$lang1[data$exp_ID == 625900] <- 'engels'
data$lang2[data$exp_ID == 625900] <- 'duits'
data$proficiency2 <- 2

##804688
data$lang1[data$exp_ID == 804688] <- 'engels'
data$lang2[data$exp_ID == 804688] <- 'duits'
data$proficiency2[data$exp_ID == 804688] <- 1 # Note, een beetje
data$lang3[data$exp_ID == 804688] <- 'italiaans'


# English, englisch, Englisch, Englisch, Enlisch, english, englisch
data$lang1 <- gsub("Nederlands |Nederlands|Nedetlands|Nederlandes|NEderlands|Nederland|Nederlands (moedertaal)", "nederlands", data$lang1)
data$lang1 <- gsub("Engels |Engels|engels ", "engels", data$lang1)
data$lang1 <- gsub("Servo-Kroatisch ", "servo-kroatisch", data$lang1)

# lower case
data$lang1 <- tolower(data$lang1)

#lang2
unique(data$lang2)
data$lang2 <- gsub("Duits|duits|Duits ", "duits", data$lang2)
data$lang2 <- gsub("Frans |Frans", "frans", data$lang2)
# lower case
data$lang2 <- tolower(data$lang2)

#lang3
unique(data$lang3)
data$lang3 <- tolower(data$lang3)
data$lang3 <- gsub("spaans ", "spaans", data$lang3)
data$lang3 <- gsub("duits ", "duits", data$lang3)

#lang4
unique(data$lang4)
# lower case
data$lang4 <- tolower(data$lang4)

#lang4
unique(data$lang5)
# lower case
data$lang5 <- tolower(data$lang5)
```

How many words were finished by participants?
```{r amount of words}

# Calculate the number of words rated by each participant
df_word_count <- data %>%
  group_by(exp_ID) %>%
  summarize(word_count = n())

summary(df_word_count)

## all participants finished at least 1 third - the same rule applied for German data
```

Duration of trial/rating and missing end times
```{r duration, UNFINISHED}

## QUESTION what is the reaction time, or minimum thinkable time for process the word?

### Participants without end time

# participant 89227, finished

## start time: 123.773
## 1st block start: 131.918
## 2nd block start: 228.318
## 3rd block start: 312.068

a <- 312.068-228.318 #duration of the 2nd block
b <- 228.318-131.918 # duration of the 1st block
c <- c(a,b) 
d <- mean(c) # mean of the duration of a block 
e <- d/20 # duration per word in the block
f <- 312.068 + d # time of start of the 3rd block + mean duration = end of the trial as such

# e = 4.50375 (duration per word/rating)
# end time = 402.143 

data$end_time[data$participant == 89227] <- 402.143 

# participant 122364, finished

a <- 620.892-515.534
b <- 515.534-396.372
c <- c(a,b)
d <- mean(c)
e <- d/20
f <- 620.892 + d 

# e = 5.613
# end time = 733.152

data$end_time[data$participant == 122364] <- 733.152 

# participant 248410, finished

a <- 310.126-230.957
b <- 230.957-145.253
c <- c(a,b)
d <- mean(c)
e <- d/20
f <- 310.126 + d

# e = 4.121825
# end time = 392.5625

data$end_time[data$participant == 248410] <- 392.5625 

# participant 366882, one modality + 2

a <- 322.981-162.071
e <- a/20
f <- e*22
g <- 162.071 + f # end time

# e = 8.0455
# end time: 339.072

data$end_time[data$participant == 366882] <- 339.072 


# participant 479477, one modality + 2 (22)

a <- 421.952-262.606
e <- a/20
f <- e*22
g <- 262.606 + f

# e = 7.9673
## end time = 437.8866

data$end_time[data$participant == 479477] <- 437.8866 


# participant 536733, 1 modality + 17

a <- 433.13-239.074
e <- a/20
f <- e*37
g <- 239.074 + f


# e = 9.7028
# end time = 598.0776

data$end_time[data$participant == 536733] <- 598.0776 


# participant 568428, finished

a <- 282.694-215.866
b <- 215.866-137.504
c <- c(a,b)
d <- mean(c)
e <- d/20
f <- 282.694 + d

# e = 3.62975
# end time = 355.289

data$end_time[data$participant == 568428] <- 355.289 


# participant 791441, finished

a <- 270.113-216.670
b <- 216.670-134.801
c <- c(a,b)
d <- mean(c)
e <- d/20
f <- 270.113 + d

# e = 3.3828
# end time = 337.769

data$end_time[data$participant == 791441] <- 337.769 


# participant 965452, finished

a <- 531.396-419.111
b <- 419.111-279.798
c <- c(a,b)
d <- mean(c)
e <- d/20
f <- 531.396 + d

# e = 6.28995
# end time = 657.195

data$end_time[data$participant == 965452] <- 657.195 


## add duration column

data$duration <- NA 
data$duration <- data$end_time - data$trial_start_time


# av duration per rating (dur / number of rows)

df_duration <- data %>% group_by(participant) %>% 
  summarize(duration_per_word = duration / n())

df_duration <- unique(df_duration)


```
Let's correct sensory preference columns
```{r sensory preference re-do}

# 1 - hearing only
# 2 - visual only
# 3 - taste only
# 4 - other 

# 5 - hearing + visual
# 6 - hearing + taste 
# 7 - hearing + other
# 8 - visual + taste ### NONE
# 9 - visual + other
# 10 - taste + other ### NONE

# 11 - hearing + visual + taste 
# 12 - hearing + taste + other ### NONE
# 13 - visual + taste + other ### NONE

# 14 - hearing + visual + taste + other - in other words all


data$sensorypref <- NA
data$sensorypref <- ifelse(!is.na(data$sensorypref1) & is.na(data$sensorypref2) & is.na(data$sensorypref3) & is.na(data$sensorypref4), '1',
                           ifelse(!is.na(data$sensorypref2) & is.na(data$sensorypref1) & is.na(data$sensorypref3) & is.na(data$sensorypref4), '2',
                                  ifelse(!is.na(data$sensorypref3) & is.na(data$sensorypref1) & is.na(data$sensorypref2) & is.na(data$sensorypref4), '3', 
                                         ifelse(!is.na(data$sensorypref4) & is.na(data$sensorypref1) & is.na(data$sensorypref2) & is.na(data$sensorypref3), '4', 
                                                ifelse(!is.na(data$sensorypref1) & !is.na(data$sensorypref2) & is.na(data$sensorypref3) & is.na(data$sensorypref4), '5',
                                                       ifelse(!is.na(data$sensorypref1) & !is.na(data$sensorypref3) & is.na(data$sensorypref2) & is.na(data$sensorypref4), '6',
                                                              ifelse(!is.na(data$sensorypref1) & !is.na(data$sensorypref4) & is.na(data$sensorypref2) & is.na(data$sensorypref3), '7',
                                                                     ifelse(!is.na(data$sensorypref2) & !is.na(data$sensorypref3) & is.na(data$sensorypref1) & is.na(data$sensorypref4), '8',
                                                                            ifelse(!is.na(data$sensorypref2) & !is.na(data$sensorypref4) & is.na(data$sensorypref1) & is.na(data$sensorypref3), '9',
                                                                                   ifelse(!is.na(data$sensorypref3) & !is.na(data$sensorypref4) & is.na(data$sensorypref1) & is.na(data$sensorypref2), '10',
                                                                                          ifelse(!is.na(data$sensorypref1) & !is.na(data$sensorypref2) & !is.na(data$sensorypref3) & is.na(data$sensorypref4), '11',
                                                                                                 ifelse(!is.na(data$sensorypref1) & !is.na(data$sensorypref3) & !is.na(data$sensorypref4) & is.na(data$sensorypref2), '12',
                                                                                                        ifelse(!is.na(data$sensorypref2) & !is.na(data$sensorypref3) & !is.na(data$sensorypref4) & is.na(data$sensorypref1), '13', '14')))))))))))))

```


Add part of speech information
```{r TODO}

# add part of speech info etc
# load concept excel with info about PoS and SemCat
concepts <- read_excel("../Data/conceptlist_info.xlsx")

# rename col Dutch as word so we can match
colnames(concepts)[7] <- "word"

# create df that matches info from df data and concepts
df_expressibility <- merge(data, concepts, by = "word")
```

Reorder columns and save
```{r reorder and save}
df_expressibility <- df_expressibility %>% select(word, modality, expressibility, SemanticCat, PoS, response_time, trial_start_time, end_time, duration, exp_ID, age, gender, natlang1, natlang2, natlang3, dialect, lang1, proficiency1, lang2, proficiency2, lang3, proficiency3, lang4, proficiency4, lang5, proficiency5, country, education, dyslexia, mobility, currentloc, sensorypref, file)

### remove duplicate rows
df_expressibility <- unique(df_expressibility)

# this is the final dataset
write.csv(df_expressibility, paste0(datasets, "df_expressibility_dutch_final.csv"), row.names = FALSE)

```

Sanity check for how many words we have rated
```{r number of ratings per word and modality}

df_expressibility <- read_csv("../Datasets/df_expressibility_final.csv")
df_ratings <- df_expressibility %>% group_by(word, modality) %>% 
  summarize(rating_amount = length(expressibility))

```

